\sys: Controllable Program Regeneration

### 2025-06-10 paper/draft.tex
Add notes to draft
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>
Refine evaluation goals
Signed-off-by: Evangelos Lamprou <vagos@lamprou.xyz>

@@

We present \sys, a system for scalable program regeneration.
\sys allows users
to express transformation goals declaratively, as logical constraints over
program properties---such as input/output behavior, side-effect freedom, target
language, trace equivalence, or modular structure. Given a source program and
such a query, \sys extracts a minimal set of relevant properties, guided by a
logic engine and a domain-specific knowledge base, and synthesizes a new
program that satisfies the desired constraints.

@@

\begin{table}[h]
\centering
  \caption{\textbf{Benchmark summary}. 
  Benchmark programs have been sourced from several sources.
  The $N$ column shows the number of programs/functions from each benchmar.
  }
\begin{tabular}{llrl}
\toprule
Benchmark                                               & Description                    & $N$ & Source \\
\midrule
  Rosetta Code                                          & Programming tasks              & 5 & \cite{rosettacode} \\
  npm utilities                                           & Utilities from npm             & 20 & \cite{regbench2025} \\
Python utilities                                        & Utilities from PyPi            & 20 & \cite{regbench2025} \\
  SSCA                                                    & Sneaky SSAs    & 3 & \cite{ev:eurosec:2022, es1, ohm2020backstabber} \\
  IOCCC                                                   & Obfuscated C code              & 2 & \cite{ioccc} \\
Microbenchmarks                                         &          & 3    & This paper \\
  \hspace{.5em} \ttt{flatmap-stream}                      & Secure regeneration            &      & \cite{es1}  \\
  \hspace{.5em} \ttt{Q_rsqrt}                             & C refactoring        &      & \cite{fast_inv_sqrt}  \\
  \hspace{.5em} \textsf{MusicDB}                          & Modularization           &      & \cite{codewithsadeemusicplayer} \\
\bottomrule
\end{tabular}
\label{tab:benchmarks}
\end{table}

>>>>

% Mention that:
% 1. They have shown great promise in program synthesis + 
% 2. They can not do negative reasoning well - 
% 3. Their output are often closely aligned to their inputs. They are gullible (susceptible to adversarial inputs). -
% 4. Difficult to verify correctness of outputs against certain properties. The
%    user cannot easily set requirements or say output should be within a range of
%    acceptance

@@

This paper present \sys, a system for scalable program regeneration.
\sys allows users
to express transformation goals declaratively, as logical constraints over
program properties---such as input/output behavior, side-effect freedom, target
language, trace equivalence, or modular structure. Given a source program and
such a query, \sys extracts a minimal set of relevant properties, guided by a
logic engine and a domain-specific knowledge base, and synthesizes a new
program that satisfies the desired constraints.

@@

\begin{table}[h]
\centering
  \caption{\textbf{Benchmark summary}. 
  Benchmark programs have been sourced from several sources.
  The $N$ column shows the number of programs/functions from each benchmar.
  }
\begin{tabular}{llrl}
\toprule
Benchmark                                               & Description                    & $N$ & Source \\
\midrule
  Rosetta Code                                          & Programming tasks              & 5 & \cite{rosettacode} \\
  npm utilities                                           & Utilities from npm             & 10 & \cite{regbench2025} \\
Python utilities                                        & Utilities from PyPi            & 1 & \cite{regbench2025} \\
  SSCA                                                    & Sneaky SSAs    & 3 & \cite{ev:eurosec:2022, es1, ohm2020backstabber} \\
  IOCCC                                                   & Obfuscated C code              & 2 & \cite{ioccc} \\
Microbenchmarks                                         &          & 3    & This paper \\
  \hspace{.5em} \ttt{flatmap-stream}                      & Secure regeneration            &      & \cite{es1}  \\
  \hspace{.5em} \ttt{Q_rsqrt}                             & C refactoring        &      & \cite{fast_inv_sqrt}  \\
  \hspace{.5em} \textsf{MusicDB}                          & Modularization           &      & \cite{codewithsadeemusicplayer} \\
\bottomrule
\end{tabular}
\label{tab:benchmarks}
\end{table}
